Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-06-30 code/rmarkdown/article.Rmd
+ prepare for pnas submission

---
title: "Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?"
author: |
  | Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Cynthia Riginos$^1$, Hugh P. Possingham$^1$, Richard A. Fuller$^1$
  | $^1$School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
  | $^2$School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
  | Correspondance should be addressed to jeffrey.hanson@uqconnect.edu.au
date: "`r format(Sys.time(), '%d %B %Y')`"
tags: ["conservation", "biodiversity", "AFLP", "reserve-selection"]
abstract: "Insert abstract here."
output:
  rmarkdown::pdf_document:
    toc: false
    toc_depth: 5
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: preamble-latex.tex
      after_body: postamble-latex.tex
fontsize: 11pt
documentclass: article
bibliography: references.bib
csl: reference-style.csl
---

@@

# set options
options(error=NULL)
```

<!--  title page -->
Short running title: Genetic surrogates in conservation \newline
Abstract word count: XXXXX \newline
Main text word count: XXXXX \newline
Number of references: XXXX \newline
Number of equations: XXXX \newline
Number of figures: XXXX \newline
Number of tables: XXXX \newline
Number of text boxes: XXXX \newline
\newpage

## Introduction
<!--  main text -->
Protected areas spearhead conservation efforts [@r440]. These places buffer species from anthropogenic impacts and provide places for species to persist. However, the resources available for conservation are limited. Thus protected areas need to be sited in places that achieve conservation goals for minimal cost. To ensure that the plans for protected areas (prioritisations) are cost-effective, they are often generated using decision support tools to identify near-optimal solutions. These prioritisations often contain plans for a network of multiple protected areas so they can preserve biological processes (eg. connectivity between populations). For instance, \texttt{Marxan} [@r22] uses amount-based targets to identify protected area networks that preserve a suitable amount of habitat for each species (eg. 100 km $^2$ of land occupied by the species needs to be secured in a reserve) and weights to penalise fragmented solutions (ie. the \textit{BLM} parameter). However, conservation planning exercises typically assume that all individuals in the same species are equivalent.

@@

## Methods
### Study system
We used species distribution and genomic data collected by the international IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods]. This data set contains information for 27 alpine plant species (Figure 1a). It was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25 km; Figure S1). Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals. They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; see Table 1 in @r478 for the number markers for each species). Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell.

@@

### Landscape data
We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively). We utilised the sampling grid used by the IntraBioDiv project as planning units to develop prioritisations. We used only planning units that contained at least a single individual ($n = `r nrow(grid.DF)`$) to reduce computational burden. To examine the potential effects of variation in cost (see below), we calculated the total human population density inside each planning unit [obtained at 1 km$^2$ resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 1b).

@@

### Adaptive and neutral genetic data
To investigate the effectiveness of our supposed surrogates for adaptive and neutral genetic variation, we first needed to identify which of the sampled loci--if any--were adaptive. We used two outlier detection methods to achieve this [an individual- and a population-level method as recommended by @r495]. The basic premise underpinning such methods is that neutral loci are expected to exhibit a specific level of variation among populations, and loci that deviate from this expectation should be under selection [reviewed in @r480]. The advantage of these methods--in contrast with environmental association analyses [@r489]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis. Loci identified as outliers in both analyses were treated as adaptive and the rest were treated as neutral. To avoid false-positives, we omitted loci from both outlier detection methods where the global frequency of the minor allele was $> 0.1$ and assumed they were neutral.

@@

## Prioritisation method
We used the unreliable representative and adequate prioritisation (URAP) formulation of the reserve selection problem [@r465]. This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' range (using amount-based targets) and a proportion of variation found across the species' range [using space-based targets and attribute spaces; _sensu_ @r465]. We treated the environmental and geographic surrogate variables as separate attribute spaces. We also treated the adaptive (if detected) and neutral variables as separate attribute spaces for each species. We set the demand points for each species and each attribute space using the values associated with the planning units. For every prioritisation we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritisation secured to assess its performance. Prioritisations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis. We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].

## Statistical analyses
To address the first aim--to determine if the environmental and geographic variables correlate with spatial patterns of adaptive and neutral genetic variation--we computed dissimilarity matrices between the planning units, and tested for correlations between matrices based on surrogate and genetic data. For each species, we computed two dissimilarity matrices between the planning units occupied by the species in terms of their environmental conditions (their coordinates in the environmental space) and geographic position (their coordinates in the geographic space). We then computed another dissimilarity matrix between the planning units using their values in the species' neutral genetic space. If the species was associated with adaptive loci, we also computed another dissimilarity matrix between the planning units using the species' adaptive genetic space. All dissimilarity matrices calculated using Euclidean distances.

@@

## Results
<!-- summary of methods -->
We successfully mapped the main spatial patterns of adaptive (if present) and neutral variation for each of the 27 species. The outlier detection methods detected adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` species. Generally, the two outlier detection methods gave fairly consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD). Of the species that were associated with adaptive genetic variation, only a few loci were detected as under selection (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD). After identifying the loci under selection, we used NMDS analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species. For all species in which we detected adaptive loci, only a couple of continuous dimensions were needed to sufficiently describe their patterns of adaptive genetic variation (all $K = `r unique(adaptive.k)`$; Table S2). On the other hand, a few additional continuous dimensions were occasionally needed to describe their patterns of neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; maximum $K = `r max(neutral.k)`$; Table S2). The ordinations resulting from the NMDS analyses were used to construct an adaptive (if present) and neutral genetic space for each species. The spatial distribution of these genetic spaces (Figures S29--S54) generally showed patterns of spatial auto-correlation (eg. _Cerastium uniflorum_, Figure SXX; _Dryas octopetala_, Figures SXXX)--a good sign that the environmental and geographic variables may prove effective surrogates.

## Discussion
<!--  introduction -->
We aimed to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning. Our study is the first to address this aim by measuring the proportion of genetic variation secured by prioritisations  generated for diverse set of species. Overall, we found that the spatial patterns of environmental and geographic variables correlate with those of adaptive and neutral patterns for most species. Furthermore, we found that the addition of environmental and geographic targets to the reserve selection process resulted in prioritisations that secured a greater proportion of adaptive and neutral genetic variation for most species.

## Acknowledgements
JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF has an Australian Research Council Future Fellowship. This work was supported by the Centre of Excellence for Environmental Decisions (CEED) and the Landscape Ecology and Conservation Group (LEC) at The University of Queensland.

## References

>>>>

---
title: "Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?"
author:
  - name: Jeffrey O. Hanson
    affiliation: a,1
  - name: Jonathan R. Rhodes
    affiliation: b
  - name: Cynthia Riginos
    affiliation: a
  - name: Hugh P. Possingham
    affiliation: a
  - name: Richard A. Fuller
    affiliation: a
address:
  - code: a
    address: School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
  - code: b
    address: School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
corresponding_author:
  - code: 1  
  - text: "To whom correspondence should be addresssed. E-mail: jeffrey.hanson@uqconnect.edu.au
author_contributions: |
  All authors developed the study and drafted the manuscript. J.O.H performed analyses.
abstract: |
  Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.
significance: |
  Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality. The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership. The Significance Statement appears in the paper itself and is required for all research papers.
acknowledgements: |
  JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF has an Australian Research Council Future Fellowship. This work was supported by the Centre of Excellence for Environmental Decisions (CEED) and the Landscape Ecology and Conservation Group (LEC) at The University of Queensland.
keywords:
  - conservation
  - biodiversity
  - AFLP
  - reserve-selection
  - genetic diversity
pnas_type: pnasresearcharticle
bibliography: references.bib
csl: reference-style.csl
output: rticles::pnas_article
---

@@

# set options
options(error=NULL)

# download basemap
data(countriesHigh)
countries.FPLY <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany', 'Slovakia', 'Czech Republic'
	)
,] %>% spFortify
grid.FPLY <- spFortify(grid.PLY)
grid.sub.FPLY <- spFortify(grid.sub.PLY)

# define ggplot2 color wheel function
gg_color_hue <- function(n) {
	hues <- seq(15, 375, length=n+1)
	hcl(h=hues, l=65, c=100)[1:n]
}
```

<!--  Introduction -->
Protected areas spearhead conservation efforts [@r440]. These places buffer species from anthropogenic impacts and provide places for species to persist. However, the resources available for conservation are limited. Thus protected areas need to be sited in places that achieve conservation goals for minimal cost. To ensure that the plans for protected areas (prioritisations) are cost-effective, they are often generated using decision support tools to identify near-optimal solutions. These prioritisations often contain plans for a network of multiple protected areas so they can preserve biological processes (eg. connectivity between populations). For instance, \texttt{Marxan} [@r22] uses amount-based targets to identify protected area networks that preserve a suitable amount of habitat for each species (eg. 100 km $^2$ of land occupied by the species needs to be secured in a reserve) and weights to penalise fragmented solutions (ie. the \textit{BLM} parameter). However, conservation planning exercises typically assume that all individuals in the same species are equivalent.

@@

Results {.unnumbered}
=====================

<!-- summary of methods -->
We successfully mapped the main spatial patterns of adaptive (if present) and neutral variation for each of the 27 species. The outlier detection methods detected adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` species. Generally, the two outlier detection methods gave fairly consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD). Of the species that were associated with adaptive genetic variation, only a few loci were detected as under selection (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD). After identifying the loci under selection, we used NMDS analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species. For all species in which we detected adaptive loci, only a couple of continuous dimensions were needed to sufficiently describe their patterns of adaptive genetic variation (all $K = `r unique(adaptive.k)`$; Table S2). On the other hand, a few additional continuous dimensions were occasionally needed to describe their patterns of neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; maximum $K = `r max(neutral.k)`$; Table S2). The ordinations resulting from the NMDS analyses were used to construct an adaptive (if present) and neutral genetic space for each species. The spatial distribution of these genetic spaces (Figures S29--S54) generally showed patterns of spatial auto-correlation (eg. _Cerastium uniflorum_, Figure SXX; _Dryas octopetala_, Figures SXXX)--a good sign that the environmental and geographic variables may prove effective surrogates.

```{r, echo=FALSE, results='asis'}
curr.file <- tempfile(fileext='.tex')
x <- correlation.model.summary.DF %>%
	select(-contains('raw')) %>%
	mutate(
		Environmental.corr.P=pvalString(Environmental.corr.P),
		Geographic.corr.P=pvalString(Geographic.corr.P),
		Geographic.Marginal.R2=pvalString(Geographic.Marginal.R2),
		Geographic.Conditional.R2=pvalString(Geographic.Conditional.R2),
		Environmental.Marginal.R2=pvalString(Environmental.Marginal.R2),
		Environmental.Conditional.R2=pvalString(Environmental.Conditional.R2),
		blank=rep('', length(Species)),
		Species=paste0('\\textit{', gsub('\\ ', ' ', Species), '}')) %>%
	mutate(Environmental.Chisq=sapply(round(Environmental.Chisq,2), as.character)) %>%
	mutate(Geographic.Chisq=sapply(round(Geographic.Chisq,2), as.character)) %>%
	select(Species:Environmental.corr.P, Geographic.Marginal.R2:Geographic.corr.P) %>%
	latex(file=curr.file, 
		cgroup=c('', 'Environmental', 'Geographic'), n.cgroup=c(1, 4, 4), col.just=rep('c', 9),
		digits=2, rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, here=TRUE,
		colheads=c('\\bfseries{Species}', '$MR^2$', '$CR^2$', '$\\chi^2$', '\\textit{P}', '$MR^2$', '$CR^2$', '$\\chi^2$', '\\textit{P}'),
		caption='Summary of maximum likelihood population-effects (MLPE) models. Briefly, for each species associated with adaptive loci, a model was fit to correlate dissimilarities among planning units where the species in terms of the adaptive genetic characteristics of individuals in them with dissimilarities in terms of the environmental conditions inside the planning units (denoted in the environmental column group). Similarly, for each species, a model was fit to correlate dissimilarities among the planning units in terms of the neutral genetic characteristics of indiviudals in them with dissimilarities in terms of the geographic position of the planning units (denoted in the geographic column group). Data shows the marginal $R^2$ ($MR^2$) and conditional $R^2$ ($CR^2$) for each model, and also the results from comparing the models to their corresponding null models. All $\\chi^2$ tests had 1 degree of freedom.'
	)
readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
	gsub(pattern='\\cline{1-', replacement='\\cline{2-6} \\cline{8-', fixed=TRUE) %>% 
	gsub(pattern='{c}{\\textit{P}}\\tabularnewline', replacement='{c}{\\textit{P}}\\T\\B\\tabularnewline', fixed=TRUE) %>% 
	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
	cat()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. Panel (a) shows the relationship between the proportion of intra-specific adaptive genetic variation and environmental variation secured in a prioritization. Similarly, panel (b) shows the relationship between the proportion of intra-specific neutral variaiton and geograpgic variation secured in a prioritization. Panels (c) and (d) show the distribution of $R^2$ values for these models.'}
# make figure
p1 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=env.surrogacy.DF) +
	xlab('Environmental variation\nsecured (%)') + 
	ylab('Adaptive genetic\nvariation secured (%)') +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black')) +
	geom_text(x=0.05, y=0.95, label='(a)') +
	coord_cartesian(xlim=c(0, 1))
for (i in seq_along(env.surrogacy.LST))
	p1 <- p1 + stat_smooth(method='glm', method.args=list(family='binomial'),
		data=filter(env.surrogacy.DF, Species==unique(Species)[i]),
		fullrange=TRUE, se=FALSE, colour='black')

p2 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=geo.surrogacy.DF) +
	xlab('Geographic variation\nsecured (%)') + 
	ylab('Neutral genetic\nvariation secured (%)') +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black', colour='transparent')) +
	geom_text(x=0.05, y=0.95, label='(b)') +
	coord_cartesian(xlim=c(0, 1))
for (i in seq_along(geo.surrogacy.LST))
	p2 <- p2 + stat_smooth(method='glm', method.args=list(family='binomial'),
		data=filter(geo.surrogacy.DF, Species==unique(Species)[i]),
		fullrange=TRUE, se=FALSE, colour='black')

p1.hist <- ggplot(aes(r2CU),
		data=filter(env.surrogate.spp.DF, !is.na(r2CU))) +
	geom_histogram(fill='grey10', bins=5) +
	ylab('\nFrequency') + xlab(expression(Environmental~variables~(R^2))) +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black'),
		plot.background=element_rect(fill='transparent', colour='transparent')) +
		geom_text(x=0.05, y=6, label='(c)') +
		coord_cartesian(xlim=c(0, 1))

p2.hist <- ggplot(aes(r2CU),
		data=filter(geo.surrogate.spp.DF, !is.na(r2CU))) +
	geom_histogram(fill='grey10', bins=10) +
	ylab('\nFrequency') + xlab(expression(Geographic~variables~(R^2))) +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black'), 
		plot.background=element_rect(fill='transparent')) +
		geom_text(x=0.05, y=8, label='(d)') +
		coord_cartesian(xlim=c(0, 1))

# render figure
grid.arrange(p1,p2,p1.hist,p2.hist,nrow=2)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=6.5, fig.cap='Planning unit selection frequencies. The first row of panels (a--c) show single-species prioritizations generated assuming equal costs. The second row of panels (d--f) show single-species prioritizations generated using opportunity costs. The third row of panels (g--i) show multi-species prioritizations generated assuming equal costs. The fourth row of panels (j--l) show multi-species generated using opportunity costs. The left column of panels (a, d, g, j) show prioritizations generated using only amount-based targets. The middle column of panels (b, e, h, k) show prioritizations generated using amount- and surrogate-based targets. The right column of panels (c, f, i, l) show prioritizations generated using amount- and genetic-based targets. The color of each planning unit indicates its selection frequency. Note that the selection frequencies for the single-species prioritizations (panels a--f) are based on their average frequency among multiple prioritizations generated for each species. The selection frequences for the multi-species prioritizations (pabels g--l) are based on the optimal pioritization (thus they are all either zero or one).'}
# prepare data for plotting
grid.sub.FPLY <- grid.sub.PLY
for (i in seq_along(single.spp.prioritisations[[1]]))
	grid.sub.FPLY@data[[paste0('Single-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations, function(x){colMeans(x[[i]]@results@selections)}))
for (i in seq_along(single.spp.prioritisations.with.cost[[1]]))
	grid.sub.FPLY@data[[paste0('Single-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- rowMeans(sapply(single.spp.prioritisations.with.cost, function(x){colMeans(x[[i]]@results@selections)}))
for (i in seq_along(multi.spp.prioritisations))
	grid.sub.FPLY@data[[paste0('Multi-species (equal costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations[[i]]@results@selections)
for (i in seq_along(multi.spp.prioritisations.with.cost))
	grid.sub.FPLY@data[[paste0('Multi-species (opportunity costs).', c('Amount', 'Surrogate', 'Genetic')[i])]] <- colMeans(multi.spp.prioritisations.with.cost[[i]]@results@selections)
grid.sub.FPLY <- spFortify(grid.sub.FPLY)
grid.sub.FPLY <- grid.sub.FPLY %>% gather(
		Combination,
		`Selection.Frequency`,
		`Single-species (equal costs).Amount`:`Multi-species (opportunity costs).Genetic`
	) %>% mutate(
		Context=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 1),
		Prioritization=sapply(strsplit(as.character(Combination), '.', fixed=TRUE), `[[`, 2)
	) %>% mutate(
		Prioritization=revalue(Prioritization, c('Amount'='Amount targets', 'Surrogate'='Amount & surrogate\ntargets', 'Genetic'='Amount & genetic\ntargets')),
		Prioritization=factor(as.character(Prioritization), levels=c('Amount targets', 'Amount & surrogate\ntargets', 'Amount & genetic\ntargets')),
		Context=factor(
			gsub(' (', '\n(', Context, fixed=TRUE),
			levels=c("Single-species\n(equal costs)", "Single-species\n(opportunity costs)",
				"Multi-species\n(equal costs)", "Multi-species\n(opportunity costs)"))
	)
panel.DF <- grid.sub.FPLY %>% select(Context, Prioritization, long, lat) %>% 
	mutate(
		lat=min(lat)+(diff(range(lat))*0.98),
		long=min(long)+(diff(range(long))*0.02)
	) %>%
	group_by(Context,Prioritization) %>% filter(c(TRUE, rep(FALSE, length(Context)-1))) %>%
	mutate(
		panel.letter=paste0('(',letters[(as.integer(Context)-1)*3 + as.integer(Prioritization)],')'),
		panel.letter=c(as.character(panel.letter[1]), rep(NA_character_, length(Context)-1))
	)
# make maps
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey85', color='grey70') +
	geom_polygon(data=grid.sub.FPLY, aes_string(x='long', y='lat', 
		group='group', fill='Selection.Frequency'),
		alpha=0.8, color='grey10') +
	geom_text(aes_string(x='long',y='lat', label='panel.letter'), data=panel.DF) +
	theme_classic() +
	theme(axis.ticks=element_blank(), axis.text=element_blank(),
		plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank(),
		strip.background = element_rect(fill='grey20'),
		strip.text = element_text(color='white'),
		legend.position='bottom', legend.key.width=unit(1.8, 'cm'),
		legend.title=element_text(vjust=-5),
		legend.margin=unit(c(0.1), 'cm'),
		axis.title=element_blank()) +
	coord_cartesian(
		xlim=buffered.range(grid.sub.FPLY$long, 0),
		ylim=buffered.range(grid.sub.FPLY$lat, 0)
	) +
	facet_grid(Context ~ Prioritization) +
	scale_fill_gradientn(name='Selection Frequency (%)',
		colors=brewer.pal(9, 'PuBu'),
		guide=guide_colorbar(title.position='bottom', title.hjust=0.5,
			ticks=element_line(color='black'), border=element_line(color='black'))
	)
```

Discussion {.unnumbered}
========================

<!--  introduction -->
We aimed to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning. Our study is the first to address this aim by measuring the proportion of genetic variation secured by prioritisations  generated for diverse set of species. Overall, we found that the spatial patterns of environmental and geographic variables correlate with those of adaptive and neutral patterns for most species. Furthermore, we found that the addition of environmental and geographic targets to the reserve selection process resulted in prioritisations that secured a greater proportion of adaptive and neutral genetic variation for most species.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=7.4, fig.cap='The effectiveness of prioritizations at securing genetic variation. The top row of panels (a--b) show the performance of single-species prioritizations generated assuming equal costs among planning costs. The middle row of panels (c--d) show the performance of multi-species prioritizations generated assuming equal costs among planning units. The bottom row of panels (e--f) show the performance of multi-species prioritizations generated using opportunity costs. The left column of panels (a, c, d) and the right column of panels (b, d, f) show the proportion of adaptive and neutral genetic variation secured (respectively). Box plots show the effectiveness of prioritizations generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species.'}
# prepare data for plotting
scenario.PDF <- scenario.DF %>% group_by(Prioritisation, Context, Metric) %>% 
	mutate(
		letter.height=max(genetic.held, na.rm=TRUE)+0.1,
		Combination=paste0(Prioritisation,'.',Metric,'.',Context)) %>%
	filter(c(TRUE, rep(FALSE, length(Context)-1))) %>% 
	ungroup() %>%
	group_by(Context, Metric) %>% 
	mutate(
		panel.letter=paste0(
			'(',letters[(as.integer(Context)*2)-1 + (as.integer(Metric)-1)],')'
		),
		panel.letter=c(as.character(panel.letter[1]), rep(NA_character_, length(Context)-1))
	) %>% 
	ungroup() %>% 
	data.frame()

scenario.SDF <- scenario.DF %>% 
	group_by(Prioritisation, Context, Metric, Species) %>% 
	summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
	data.frame()

# make plot
suppressWarnings({
	ggplot(aes(x=Prioritisation,y=genetic.held),
		data=scenario.SDF) +
		geom_boxplot(position=position_dodge(0.9)) +
		geom_text(aes(x=0.6, y=1.15, label=panel.letter),
			data=scenario.PDF) +
		scale_fill_manual(name='Prioritisation',
			values=c('grey80','grey50','grey20')) +
		geom_abline(slope=0, color='red', linetype='dashed',
			intercept=rapr.params.LST[[MODE]]$scenario.analysis$genetic.target) +
		ylab('Genetic variation secured (%)') +
		xlab('') +
		scale_y_continuous(breaks=c(0,0.25,0.5,0.75,1), limits=c(NA, 1.2)) +
		theme_classic() +
		theme(strip.background = element_rect(fill='grey20'),
			strip.text = element_text(color='white'),
			panel.border = element_rect(color='black', fill=NA),
			axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
		facet_grid(Context ~ Metric)
})
```

Materials and Methods {.unnumbered}
===================================

Study system {#study-system .unnumbered}
----------------------------------------

We used species distribution and genomic data collected by the international IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods]. This data set contains information for 27 alpine plant species (Figure 1a). It was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25 km; Figure S1). Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals. They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; see Table 1 in @r478 for the number markers for each species). Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell.

```{r, echo=FALSE, fig.width=6, fig.height=1.5, fig.cap="Map of the study area. Squares denote planning units. Panel (a) depicts patterns of species riches. Planning units with a brighter color are inhabited by more species. Panel (b) shows the distribution of opportunity cost across the study area (computed as total population density). Note that a quantile-based color ramp is used in this panel."}

# calculate species richness
grid.PLY$Species_richness <- grid.PLY@data %>%
	select(5:(4+n_distinct(spp.samples.DF$species))) %>% as.matrix() %>% rowSums()

panel.DF <- grid.sub.FPLY %>% 
	summarize(lat=min(lat)+(diff(range(lat))*0.96),
		long=min(long)+(diff(range(long))*0.04))

# plot species richness
p1 <- ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey85', color='grey70') +
	geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat, 
		group=group, fill=Species_richness), color='grey10') +
	scale_fill_gradientn(name='Richness',
		colors=brewer.pal(9,'PuBu'),
		guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
	) +
	theme_classic() +
	theme(
		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
		panel.border=element_rect(color='black', fill=NA, size=1)
	) +
	coord_cartesian(
		xlim=buffered.range(grid.FPLY$long, 0),
		ylim=buffered.range(grid.FPLY$lat, 0)
	) + geom_text(aes_string(x='long',y='lat'), data=panel.DF, label='(a)')

qn01 <- scales::rescale(quantile(grid.PLY@data$pop.density, seq(0,1,length.out=9), na.rm=TRUE))
p2 <- ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey85', color='grey70') +
	geom_polygon(data=spFortify(grid.PLY), aes(x=long, y=lat, 
		group=group, fill=pop.density), color='grey10') +
	scale_fill_gradientn(name='Cost',
		colors=brewer.pal(9, 'PuBu'),
		values=qn01,
		guide=guide_colorbar(ticks=element_line(color='black'), border=element_line(color='black'))
	) +
	theme_classic() +
	theme(
		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"), axis.title=element_blank(),
		panel.border=element_rect(color='black', fill=NA, size=1)
	) +
	coord_cartesian(
		xlim=buffered.range(grid.FPLY$long, 0),
		ylim=buffered.range(grid.FPLY$lat, 0)
	)  + geom_text(aes_string(x='long',y='lat'), data=panel.DF, label='(b)')

grid.arrange(p1,p2,nrow=1)
```

@@

Landscape data {#landscape-data .unnumbered}
--------------------------------------------

We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively). We utilised the sampling grid used by the IntraBioDiv project as planning units to develop prioritisations. We used only planning units that contained at least a single individual ($n = `r nrow(grid.DF)`$) to reduce computational burden. To examine the potential effects of variation in cost (see below), we calculated the total human population density inside each planning unit [obtained at 1 km$^2$ resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 1b).

@@

Adaptive and neutral genetic data {#adaptive-and-neutral-genetic-data .unnumbered}
----------------------------------------------------------------------------------

To investigate the effectiveness of our supposed surrogates for adaptive and neutral genetic variation, we first needed to identify which of the sampled loci--if any--were adaptive. We used two outlier detection methods to achieve this [an individual- and a population-level method as recommended by @r495]. The basic premise underpinning such methods is that neutral loci are expected to exhibit a specific level of variation among populations, and loci that deviate from this expectation should be under selection [reviewed in @r480]. The advantage of these methods--in contrast with environmental association analyses [@r489]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis. Loci identified as outliers in both analyses were treated as adaptive and the rest were treated as neutral. To avoid false-positives, we omitted loci from both outlier detection methods where the global frequency of the minor allele was $> 0.1$ and assumed they were neutral.

@@

Prioritisation method {#prioritisation-method .unnumbered}
----------------------------------------------------------

We used the unreliable representative and adequate prioritisation (URAP) formulation of the reserve selection problem [@r465]. This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' range (using amount-based targets) and a proportion of variation found across the species' range [using space-based targets and attribute spaces; _sensu_ @r465]. We treated the environmental and geographic surrogate variables as separate attribute spaces. We also treated the adaptive (if detected) and neutral variables as separate attribute spaces for each species. We set the demand points for each species and each attribute space using the values associated with the planning units. For every prioritisation we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritisation secured to assess its performance. Prioritisations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis. We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].

Statistical analyses {#statistical-analyses .unnumbered}
----------------------------------------------------------

To address the first aim--to determine if the environmental and geographic variables correlate with spatial patterns of adaptive and neutral genetic variation--we computed dissimilarity matrices between the planning units, and tested for correlations between matrices based on surrogate and genetic data. For each species, we computed two dissimilarity matrices between the planning units occupied by the species in terms of their environmental conditions (their coordinates in the environmental space) and geographic position (their coordinates in the geographic space). We then computed another dissimilarity matrix between the planning units using their values in the species' neutral genetic space. If the species was associated with adaptive loci, we also computed another dissimilarity matrix between the planning units using the species' adaptive genetic space. All dissimilarity matrices calculated using Euclidean distances.

@@

\showmatmethods
\showacknow
\pnasbreak
